---
title: "SEER Thyroid"
author: "Dan Weinberger"
date: '2023-07-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(lubridate)
library(MASS)
library(ggplot2)
library(patchwork) # combining plots

source('./R/RunControl.R')
source('./R/OutputsControl.R')

```

## Read in Data

```{r}
a1 <- read_excel('./CONFIDENTIAL/NEW.SEERThy.24.4.23.xlsx', sheet='Rate x hist x mo', skip=1) %>%
  mutate( date = as.Date(paste(Year, Month, '01', sep='-'), '%Y-%B-%d')) %>%
  rename(monthdays = 'd/mo', Outcome=Count) %>%
  dplyr::select( Histology,date, Outcome, Population,monthdays ) %>%
    mutate(log_offset= log(Population/monthdays*30.3)) %>%
  arrange(Histology, date) 


#CHECK THIS--WEIRD MISSINGNESS PATTERN
a2 <- read_excel('./CONFIDENTIAL/NEW.SEERThy.24.4.23.xlsx', sheet='Size cat x mo PAP', skip=1) %>%
  mutate( date = as.Date(paste(Yr, Mo, '01', sep='-'), '%Y-%B-%d')) %>%
  rename( Outcome=Count, Population=Pop) %>%
  mutate( Histology = paste(Hist, `Size Cat`)) %>%
  dplyr::select( Histology,date, Outcome, Population ) %>%
    mutate(hist_grp = if_else(Histology %in% c("Papillary <2cm" ), 'Pap_u_2cm',
                              if_else(Histology %in% c("Papillary >2cm" ) ,'Pap_over_2cm','other'  ))) %>%
    arrange(hist_grp, date) %>%
  group_by(hist_grp, date) %>%
  summarize(Outcome=sum(Outcome), Population=mean(Population)) %>%
   mutate( Population = if_else(date<='2016-03-01',63621556, Population), 
      monthdays= lubridate::days_in_month(date),
          log_offset= log(Population/monthdays*30.3) ) %>%
  ungroup() %>%
  filter( !is.na(date))


size1 <- read_excel('./CONFIDENTIAL/NEW.SEERThy.24.4.23.xlsx',sheet='Size_Yr') %>%
  filter(Size <989) %>%
  mutate(year_cat=as.factor(Yr),
         Hist = if_else(Hist =='Papiillary', 'Papillary', Hist)) 

```
The six panels in each figure will be: papillary cancers 2cm or less, papillary cancers >2cm, follicular, Oncocytic (formerly Hurthle cell), medullary and anaplastic.

## Panel A: Papillary <2 cm {.tabset}
```{r}
b1 <- a2 %>%
  filter(hist_grp == 'Pap_u_2cm') %>%
  arrange(date) %>%
  mutate(index=row_number())

set.start.date =as.Date('2020-03-01')

form1a <- ' ~ index + offset(log_offset)'
outcome <- 'Outcome_pre '
form1 <- paste0(outcome, form1a)

mod1 <- RunControl(ds=b1 ,set.start.date=set.start.date, set.form_cont1=form1)

out1 <-  OutputsControl(fitted.mod=mod1, niter=10000)
```

### RR trend
```{r}
out1$p.rr.trend
```
### Obs vs expected trend
```{r}
fig2a <- out1$p.preds
fig2a
```
### Cumulative prevented trend

```{r}
fig3a <- out1$p.cum_prevented
fig3a
```
## Panel B: Papillary >=2 cm {.tabset}
```{r}
b1 <- a2 %>%
  filter(hist_grp == 'Pap_over_2cm') %>%
  arrange(date) %>%
  mutate(index=row_number())

set.start.date =as.Date('2020-03-01')

form1a <- ' ~ index + offset(log_offset)'
outcome <- 'Outcome_pre '
form1 <- paste0(outcome, form1a)

mod1 <- RunControl(ds=b1 ,set.start.date=set.start.date, set.form_cont1=form1)

out2 <-  OutputsControl(fitted.mod=mod1, niter=10000)
```

### RR trend
```{r}
out2$p.rr.trend
```
### Obs vs expected trend
```{r}
fig2b <- out2$p.preds
fig2b
```
### Cumulative prevented trend

```{r}
fig3b <- out2$p.cum_prevented
fig3b
```


## Panel C: Follicular {.tabset}
```{r}
b1 <- a1 %>%
  filter(Histology == 'Follicular') %>%
  arrange(date) %>%
  mutate(index=row_number())

set.start.date =as.Date('2020-03-01')

form1a <- ' ~ index + offset(log_offset)'
outcome <- 'Outcome_pre '
form1 <- paste0(outcome, form1a)

mod1 <- RunControl(ds=b1 ,set.start.date=set.start.date, set.form_cont1=form1)

out3 <-  OutputsControl(fitted.mod=mod1, niter=10000)
```

### RR trend
```{r}
out3$p.rr.trend
```
### Obs vs expected trend
```{r}
fig2c <- out3$p.preds
fig2c
```
### Cumulative prevented trend

```{r}
fig3c <- out3$p.cum_prevented
fig3c
```


## Panel D:  Oncocytic (formerly Hurthle cell) {.tabset}
```{r}
b1 <- a1 %>%
  filter(Histology == 'Oncocytic') %>%
  arrange(date) %>%
  mutate(index=row_number())

set.start.date =as.Date('2020-03-01')

form1a <- ' ~ index + offset(log_offset)'
outcome <- 'Outcome_pre '
form1 <- paste0(outcome, form1a)

mod1 <- RunControl(ds=b1 ,set.start.date=set.start.date, set.form_cont1=form1)

out4 <-  OutputsControl(fitted.mod=mod1, niter=10000)
```

### RR trend
```{r}
out4$p.rr.trend
```
### Obs vs expected trend
```{r}
fig2d <- out4$p.preds
fig2d
```
### Cumulative prevented trend

```{r}
fig3d <- out4$p.cum_prevented
fig3d
```


## Panel E: Medullary {.tabset}
```{r}
b1 <- a1 %>%
  filter(Histology == 'Medullary') %>%
  arrange(date) %>%
  mutate(index=row_number())

set.start.date =as.Date('2020-03-01')

form1a <- ' ~ index + offset(log_offset)'
outcome <- 'Outcome_pre '
form1 <- paste0(outcome, form1a)

mod1 <- RunControl(ds=b1 ,set.start.date=set.start.date, set.form_cont1=form1)

out5 <-  OutputsControl(fitted.mod=mod1, niter=10000)
```

### RR trend
```{r}
out5$p.rr.trend
```
### Obs vs expected trend
```{r}
fig2e <- out5$p.preds
fig2e
```
### Cumulative prevented trend

```{r}
fig3e <- out5$p.cum_prevented
fig3e
```

## Panel F: Anaplastic {.tabset}
```{r}
b1 <- a1 %>%
  filter(Histology == 'Anaplastic') %>%
  arrange(date) %>%
  mutate(index=row_number())

set.start.date =as.Date('2020-03-01')

form1a <- ' ~ index + offset(log_offset)'
outcome <- 'Outcome_pre '
form1 <- paste0(outcome, form1a)

mod1 <- RunControl(ds=b1 ,set.start.date=set.start.date, set.form_cont1=form1)

out6 <-  OutputsControl(fitted.mod=mod1, niter=10000)
```

### RR trend
```{r}
out6$p.rr.trend
```
### Obs vs expected trend
```{r}
fig2f <- out6$p.preds
fig2f
```
### Cumulative prevented trend

```{r}
fig3f <- out6$p.cum_prevented
fig3f
```
papillary cancers 2cm or less, papillary cancers >2cm, follicular, Oncocytic (formerly Hurthle cell), medullary and anaplastic.

```{r, fig.width=7, fig.height=4}
(
  (fig2a + ylim(0,800) +ggtitle('Papillary <2cm'))+
  (fig2b + ylim(0,800)+ggtitle('Papillary \u2265 2cm'))+
  (fig2c + ylim(0,150) + ggtitle('Follicular')) 
  )/ (
  (fig2d + ylim(0,150)+ ggtitle('Oncocytic (Hurthle cell'))+
  (fig2e + ylim(0,150)+ ggtitle('Medullary'))+
  (fig2f + ylim(0,150)+ ggtitle('Anaplastic')) 
  )+
   plot_annotation(tag_levels = 'A')

```
log
```{r, fig.width=7, fig.height=4}
(
  (out1$p.preds.log + ylim(1,7) +ggtitle('Papillary <2cm'))+
  (out2$p.preds.log  + ylim(1,7)+ggtitle('Papillary \u2265 2cm'))+
  (out3$p.preds.log  + ylim(1,7) + ggtitle('Follicular')) 
  )/ (
  (out4$p.preds.log  + ylim(1,7)+ ggtitle('Oncocytic (Hurthle cell'))+
  (out5$p.preds.log  + ylim(1,7)+ ggtitle('Medullary'))+
  (out6$p.preds.log  + ylim(1,7)+ ggtitle('Anaplastic')) 
  )+
   plot_annotation(tag_levels = 'A')

```

```{r, fig.width=6, fig.height=4}
(
  (fig3a + ylim(-100,1500) +ggtitle('Papillary <2cm'))+
  (fig3b + ylim(-100,1500)+ggtitle('Papillary \u2265 2cm'))+
  (fig3c + coord_cartesian(ylim = c(-40, 300))+ ggtitle('Follicular'))  
  )/ (
  (fig3d +coord_cartesian(ylim = c(-40, 300))+ ggtitle('Oncocytic'))+
  (fig3e + coord_cartesian(ylim = c(-40, 300))+ ggtitle('Medullary'))+
  (fig3f + coord_cartesian(ylim = c(-40, 300))+ ggtitle('Anaplastic')) 
  )+
   plot_annotation(tag_levels = 'A')

```

```{r}
all.results <- list(out1,out2, out3, out4, out5, out6)

cbind( c('Papillary <2cm', 'Papillary >=2cm', 'Follicular', 'Oncocytic', 'Medullary', 'Anaplastic') , round(t(sapply(all.results, '[[', 'rr.q.post')),2))

```

```{r}
all.results <- list(out1,out2, out3, out4, out5, out6)

cbind.data.frame( 'Type'=c('Papillary <2cm', 'Papillary >=2cm', 'Follicular', 'Oncocytic', 'Medullary', 'Anaplastic') , round(t(sapply(all.results, '[[', 'cum.post.t.q.last')),-1)  ) 

```

## Size regression
```{r}

hists <- unique(size1$Hist)

for(i in 1:length(hists)){
    
print(hists[i])
    
ana <- size1 %>%
  filter(Hist==hists[i])

mod1 <- lm(Size ~ year_cat, data=ana)

print(summary(mod1))
}


```

